\section{Visuomotor Framework}
\label{sec:method}

In this section, we formally describe the architectures and representations employed in our study of bimanual garment folding. Our framework, illustrated in the system architecture diagram in Figure \ref{fig:system_diagram}, integrates generative temporal modeling with high-capacity visual backbones and mid-level geometric priors.

\subsection{Action Chunking with Transformers (ACT)}
The primary architecture used for fine-grained bimanual manipulation is Action Chunking with Transformers (ACT) \cite{zhao2023act}. ACT addresses the "compounding error" problem in behavioral cloning by predicting a sequence (chunk) of actions $a_{t:t+k}$ given the current observation $o_t$. 

The model consists of a Conditional Variational Autoencoder (CVAE) and a Transformer decoder. The CVAE encoder maps a sequence of actions to a latent variable $z$, which represents the "style" or "intent" of the trajectory. During inference, the latent $z$ is set to its mean value. The Transformer decoder then takes the observation $o_t$ and the latent $z$ to predict the action chunk $\hat{a}_{t:t+k}$. 

In our experiments, we utilize a default chunk size of $k=100$. By default, the model operates in an \textbf{open-loop} fashion, where the number of action steps executed per inference ($n_{action}$) is equal to the chunk size ($n_{action} = k$). However, for our most robust configurations, we employ a \textbf{closed-loop} reactive protocol with $n_{action} = 1$, allowing the policy to re-plan at every timestep based on updated visual feedback. This formulation inherently captures the complex temporal dependencies and high-frequency coordination required for bimanual manipulation of deformable objects.

\subsection{Vision-Language-Action (VLA) Architectures}
We evaluate several Vision-Language-Action models, including SmolVLA \cite{smolvla2025}, $\pi_{0.5}$ \cite{pi05_2026}, and X-VLA \cite{zheng2025xvlasoftpromptedtransformerscalable}. These models utilize large-scale pre-trained Vision-Language Models (VLMs) as visual encoders, which are integrated with action-prediction heads. To adapt these massive backbones to the specialized domain of garment folding, we employ Parameter-Efficient Fine-Tuning (PEFT) via Low-Rank Adaptation (LoRA) \cite{hu2021lora}. LoRA allows for the optimization of a small subset of parameters by introducing low-rank decomposition matrices into the attention layers, thereby preventing catastrophic forgetting while enabling task-specific alignment.

\begin{table*}[t]
\centering
\caption{Per-garment success rates (\%) across different architectures. T-L: Top Long, T-S: Top Short, P-L: Pant Long, P-S: Pant Short. Macro-Avg is the mean across all categories. Leaderboard is the official evaluation on 20 garments per category (80 total). Hyphen (-) indicates that the model was not evaluated for that category.}
\label{tab:per_garment_results}
\resizebox{\textwidth}{!}{%
\begin{tabular}{lcccccc}
\toprule
\textbf{Model Configuration} & \textbf{T-L (\%)} & \textbf{T-S (\%)} & \textbf{P-L (\%)} & \textbf{P-S (\%)} & \textbf{Macro-Avg (\%)} & \textbf{Leaderboard (\%)} \\
\midrule
Diffusion Policy (DP) \cite{chi2023diffusionpolicy}$^\ast$ & 5.00\% & 0.00\% & 3.00\% & 38.33\% & 11.67\% & - \\
SmolVLA \cite{smolvla2025} & 56.67\% & 11.67\% & 35.00\% & 73.33\% & 44.17\% & 34.50\% \\
SmolVLA + Augmentations & 55.00\% & 20.00\% & 30.00\% & 81.67\% & 46.67\% & - \\
SmolVLA + Aug + PEFT (LoRA) & 65.00\% & 28.33\% & 43.00\% & 78.33\% & 53.75\% & 49.50\% \\
X-VLA \cite{zheng2025xvlasoftpromptedtransformerscalable} + PEFT & - & 0.00\% & - & - & 0.00\% & - \\
$\pi_{0.5}$ \cite{pi05_2026} (Base) & - & 0.00\% & - & - & 0.00\% & - \\
$\pi_{0.5}$ \cite{pi05_2026} + PEFT & - & 3.33\% & - & - & 0.83\% & - \\
\midrule
ACT (Base) \cite{zhao2023act} & 75.00\% & 45.00\% & 38.00\% & 78.33\% & 58.96\% & - \\
ACT + Depth & 70.83\% & 50.00\% & 47.00\% & 77.50\% & 61.25\% & - \\
ACT + Depth + Aug (Chunk 64) & - & 40.00\% & 29.00\% & - & 34.59\% & - \\
ACT + Depth + Aug (Chunk 100) & 61.67\% & 43.33\% & 50.00\% & 84.17\% & 59.79\% & - \\
ACT + Depth + Aug (Chunk 128) & 65.00\% & 38.33\% & 35.00\% & 79.17\% & 54.38\% & - \\
\textbf{ACT + Depth + Aug + $n_{action}=1$}$^\dagger$ & \textbf{66.67\%} & \textbf{48.33\%} & \textbf{53.00\%} & \textbf{83.33\%} & \textbf{62.92\%} & \textbf{Pending} \\
ACT + Depth + Aug + $n_{action}=1$ + DINOv2 & 77.50\% & 33.33\% & 57.00\% & 86.67\% & 63.54\% & - \\
\bottomrule
\multicolumn{7}{l}{$^\ast$ Evaluated on 5 episodes per garment category; all other models evaluated on 10 episodes.} \\
\multicolumn{7}{l}{$^\dagger$ Denotes official final submission to the LeHome Challenge leaderboard.}
\end{tabular}%
}
\end{table*}

\subsection{Dataset and Evaluation Protocol}
The study utilizes the LeHome Challenge dataset \cite{lehome2026}, which comprises 12 unique garments per category (long-sleeved tops, short-sleeved tops, long pants, and shorts), for a total of 48 initial garments provided by the organizers. Our training pipeline utilizes 10 garments from each category, while the remaining 2 garments per category are held out for local unseen evaluation. 

The final performance is validated through two distinct protocols:
\begin{itemize}[leftmargin=*,noitemsep]
    \item \textbf{Local Evaluation}: Conducted on the 10 seen and 2 unseen garments per category provided in the simulation environment.
    \item \textbf{Leaderboard Evaluation}: The models are submitted for evaluation on a comprehensive set of 20 unique garments per category. This includes the 12 original garments plus 8 completely secret garments per category, providing a rigorous test of "true" generalization across 80 total garment variations.
\end{itemize}


\subsection{Visual Representations and Backbones}
Our framework leverages diverse visual backbones and mid-level representations to reduce appearance-based overfitting:
\begin{itemize}[leftmargin=*,noitemsep]
    \item \textbf{DINOv2}: We utilize DINOv2 \cite{oquab2024dinov2learningrobustvisual} as a high-capacity visual backbone. DINOv2 provides dense, robust features learned through self-supervised training on large-scale datasets, which prove critical for reasoning about the complex geometries of textiles.
    \item \textbf{Colorized Depth}: To provide explicit geometric priors, we augment RGB observations with colorized depth maps. Raw depth values are mapped to a 3-channel JET-style gradient, allowing the visual encoders to leverage pre-trained features while perceiving the 3D structure of folds and wrinkles.
    \item \textbf{Data Augmentations}: During training, we apply a suite of stochastic augmentations, including random cropping and rotation, to both RGB and colorized depth inputs. This regularizes the model and improves generalization to novel garment styles and poses.
\end{itemize}
