\section{Experiments and Results}
\label{sec:experiments}

\subsection{Quantitative Analysis}

We evaluate our models on the four garment categories of the LeHome Challenge: long-sleeved tops (T-L), short-sleeved tops (T-S), long pants (P-L), and shorts (P-S). The evaluation uses the success rate metric based on cloth vertex alignment against a canonical state.

Table \ref{tab:per_garment_results} presents the success rates for each model across all categories. The results show a clear progression in performance as we transitioned from diffusion-based models to VLA variants and finally to optimized ACT configurations.

\begin{table*}[t]
\centering
\caption{Per-garment success rates (\%) across different architectures. T-L: Top Long, T-S: Top Short, P-L: Pant Long, P-S: Pant Short. Macro-Avg is the mean across all categories.}
\label{tab:per_garment_results}
\resizebox{\textwidth}{!}{%
\begin{tabular}{lccccc}
\toprule
\textbf{Model Strategy} & \textbf{T-L (\%)} & \textbf{T-S (\%)} & \textbf{P-L (\%)} & \textbf{P-S (\%)} & \textbf{Macro-Avg (\%)} \\
\midrule
Diffusion Policy (DP) \cite{chi2023diffusionpolicy} & 5.00\% & 0.00\% & 3.00\% & 38.33\% & 11.67\% \\
SmolVLA \cite{smolvla2025} & 56.67\% & 11.67\% & 35.00\% & 73.33\% & 44.17\% \\
SmolVLA + Augmentations & 55.00\% & 20.00\% & 30.00\% & 81.67\% & 46.67\% \\
SmolVLA + Aug + PEFT (LoRA) & 65.00\% & 28.33\% & 43.00\% & 78.33\% & 53.75\% \\
X-VLA \cite{zheng2025xvlasoftpromptedtransformerscalable} & 0.00\% & 0.00\% & 0.00\% & 0.00\% & 0.00\% \\
$\pi_{0.5}$ \cite{pi05_2026} & 0.00\% & 0.00\% & 0.00\% & 0.00\% & 0.00\% \\
$\pi_{0.5}$ + PEFT & 0.00\% & 3.33\% & 0.00\% & 0.00\% & 3.33\% \\
ACT (Base) \cite{zhao2023act} & 75.00\% & 45.00\% & 38.00\% & 78.33\% & 58.96\% \\
ACT + Depth & 70.83\% & 50.00\% & 47.00\% & 77.50\% & 61.25\% \\
ACT + Depth + Aug (Chunk 100) & 61.67\% & 43.33\% & 50.00\% & 84.17\% & 59.79\% \\
ACT + Depth + Aug (Chunk 64) & -- & 40.00\% & 29.00\% & -- & 34.59\% \\
ACT + Depth + Aug (Chunk 128) & 65.00\% & 38.33\% & 35.00\% & 79.17\% & 54.38\% \\
ACT + Depth + Aug + $n_{action}=1$ & 66.67\% & 48.33\% & 53.00\% & 83.33\% & 62.92\% \\
\textbf{ACT + Depth + Aug + $n_{action}=1$ + DINOv2} & \textbf{77.50\%} & 33.33\% & \textbf{57.00\%} & \textbf{86.67\%} & \textbf{63.54\%} \\
\bottomrule
\end{tabular}%
}
\end{table*}

\subsection{Qualitative Model Analysis}

Our analysis of the model trajectories and failure modes reveals critical insights into the behavior of each architecture:

\begin{itemize}[leftmargin=*,noitemsep]
    \item \textbf{SmolVLA Variants}: The base SmolVLA model performed significantly better than DP, achieving smooth actions in many cases. However, video analysis revealed jittery movements even during successful trials. Data augmentations slightly improved smoothness, while PEFT (training the VLM backbone) further refined gripper control and overall stability.
    \item \textbf{Large VLA Failures}: X-VLA and $\pi_{0.5}$ exhibited poor performance. X-VLA failed to even close its gripper, while $\pi_{0.5}$ struggled to localize the garment. Although PEFT improved $\pi_{0.5}$'s spatial understanding, it still failed to complete the folding sequence. We attribute these failures to the limited scale of the demonstration dataset, which was insufficient for fine-tuning such large models.
    \item \textbf{ACT Superiority}: ACT outperformed all other models, demonstrating significantly smoother trajectories and better coordination. The integration of depth priors (ACT + Depth) provided a noticeable boost by clarifying the 3D structure of the cloth.
\end{itemize}

\subsection{Sensitivity Analysis}

We investigated the impact of architectural choices and temporal parameters on ACT performance:

\begin{itemize}[leftmargin=*,noitemsep]
    \item \textbf{Chunk Size}: We compared chunk sizes of 64, 100, and 128. A chunk size of 100 provided the best results. Smaller chunks (64) struggled with long garments like pants, likely because they require more extended planning. Larger chunks (128) saw a decrease in performance across both short and long garments due to error accumulation.
    \item \textbf{Execution Steps}: Reducing the number of execution steps to 1 (\texttt{n\_action\_steps} = 1) while maintaining a chunk size of 100 improved the macro-average to 62.92\%. This suggests that more frequent re-planning allows the model to better adjust to the deformable dynamics of the fabric.
    \item \textbf{Backbone Impact}: Replacing the ResNet-18 backbone with DINOv2 resulted in a 1\% increase in the macro-average (63.54\%). While performance on short-sleeved tops decreased, it significantly improved on the other three categories, indicating better feature extraction for complex geometries.
\end{itemize}

\subsection{Generalization and Limitations}

Table \ref{tab:unseen_per_garment} shows success rates on unseen styles. While ACT + Depth achieves high success on seen garments, the addition of augmentations (ACT + Depth + Aug) results in better generalization, particularly for the pant categories.

\begin{table*}[t]
\centering
\caption{Per-garment success rates (\%) on unseen garment styles.}
\label{tab:unseen_per_garment}
\resizebox{\textwidth}{!}{%
\begin{tabular}{lccccc}
\toprule
\textbf{Model} & \textbf{T-L (\%)} & \textbf{T-S (\%)} & \textbf{P-L (\%)} & \textbf{P-S (\%)} & \textbf{Unseen Avg (\%)} \\
\midrule
SmolVLA \cite{smolvla2025} & 30.00\% & 0.00\% & 20.00\% & 60.00\% & 27.50\% \\
ACT (Base) & 85.00\% & 30.00\% & 30.00\% & 45.00\% & 47.50\% \\
ACT + Depth & 60.00\% & 35.00\% & 10.00\% & 55.00\% & 40.00\% \\
ACT + Depth + Aug + $n_{action}=1$ + DINOv2 & 50.00\% & 15.00\% & 55.00\% & 55.00\% & 43.75\% \\
\bottomrule
\end{tabular}%
}
\end{table*}

The "Top Short" category remains the most difficult across all models due to the minimal surface area available for grasping the sleeves. Infrastructure constraints, such as the lack of a GUI on the HPC environment, prevented us from using real-time human intervention to address these edge cases.
