\section{Conclusion}
\label{sec:conclusion}

In this work, we explored the application of generative imitation learning for bimanual garment folding in the context of the ICRA 2026 LeHome Challenge. Our investigation led us to shift from our initial Diffusion Policy proposal to an architecture based on Action Chunking with Transformers (ACT). We demonstrated that ACT's ability to model action sequences as chunks is significantly better suited for coordinated bimanual trajectories compared to standard Diffusion Policy or large-scale VLA models when trained on specialized datasets.

By integrating geometric priors through colorized depth and employing a DINOv2 visual backbone, we achieved a macro-average success rate of 63.54\% and demonstrated robust generalization. Our findings highlight the importance of choosing architectures that capture complex temporal dependencies and the effectiveness of mid-level geometric representations.

\subsection{Limitations}
Our study was significantly impacted by infrastructure constraints. We utilized a shared HPC environment with an NVIDIA RTX 4090 (24GB), which was often unavailable, leading to substantial delays in training and evaluation. Furthermore, the lack of a graphical interface on the HPC prevented the implementation of human-in-the-loop frameworks like DAgger, as real-time visual feedback and corrective demonstrations were not feasible.

\subsection{Future Work}
Future research will focus on overcoming these infrastructure barriers to implement online corrective feedback. We also plan to explore combining the strengths of transformer-based action chunking with the vast pre-trained knowledge of VLAs by using foundation models as visual encoders for ACT. Additionally, we aim to extend this framework to multi-task scenarios and more complex robotic embodiments, such as high-DOF dexterous hands, where the data efficiency of ACT and the semantic understanding of VLAs can be combined for more capable manipulation systems.
